Study of Higgs plus single top production using events with a same sign dimuon at the Large Hadron Collider

### 2019-11-21 tesis_hiram/Chapter1/chapter1.tex
new
correccion

In the recent years, there have been many studies in the Higgs production processes at CERN in pp collisions, such as gluon fusion (ggF), VBF, WH, ZH and $t\bar{t}$H and people have obtained consistent results in the search of Higgs in those processes. 
In this work, we will analyze the production of Higgs boson in association with a
single top quark (tH) in proton-proton collisions with the CMS experiment of the LHC. 
The motivation of the work is this mechanism of production of the Higgs boson has not been observed before by any experiment. The exploration of Higgs production on the tH channel is subject relatively new. %The SM is used to study the tH channel and compare it with other Higgs production mechanisms results.
\\
Many channels are yet out of
reach experimentally. At the same time, many theoretical
proposals are yet to be put on test. It is possible that studies on the Higgs boson give evidence of small deviations from the SM and test new physics that are beyond SM such as String Theory and Supersymmetry. \\
%********************************** % Third Section *************************************

@@

 Quarks combine to form composite particles called hadrons. Quarks cannot be found in isolation, they always exist in form of hadrons that can be mesons or baryons. \\
Mesons, are bosons made up of a quark-antiquark pair, such as pion ($\pi$) and kaon ($K$).Mesons were theorized by Yukawa in 1937 and discovered in 1947 by Powell\cite{griff}.
Baryons are fermions made of 3 quarks. Examples of baryons are protons and neutrons. Both particles can interact via strong force. Since quarks and gluons only exist in bound states, it is impossible to observe isolated quarks except the top quark, which decays before it has time to form a
bound state\cite{gross.griff}.
\\

@@

Quarks have characteristics such as electric charge, mass, color charge, and spin. The quarks up, charm and top have in common a charge of $\frac{2}{3}$, while the down, strange and bottom quarks have charge of $-\frac{1}{3}$. Among them, the top quark is the most massive of all, while the up and down are the least massive.\\ 
Some of the particles in the SM are not stable, so they decays and form particles with lower masses. Other particles do not decay or their lifetimes are very long. The final states can be represented with a Feynman diagram according to the SM vertices\cite{mark}.

@@

The last group of the family are the bosons,which have a spin 1 and 0. 
There are 5 bosons: $W^{\pm}$, $Z$ with neutral charge are responsible for weak force; photon, a massless particle, is responsible for electromagnetic force; gluon which also massless, is related to strong force and Higgs boson, recently discovered, is in charge of give the property of mass to particles via Higgs mechanism. Bosons let other particles to interact, even themselves, which is shown in the figure \ref{sm}. 
\\
The $W$,$Z$ and Higgs bosons, can interact with the most of the particles, while the gluon can only interact with quarks and itself.
Those bosons act as mediators in decays and creation of complex
particles, as mesons and baryons.\\

@@

There are experiments that detects particles around the world, such as the large hadron collider (LHC) in Europe, the Fermilab in United States of America, KEK in Japan, etc. More information about the LHC will be described in the next chapters.
\\

The table \ref{SM table} contains the characteristics of each particle in the SM. The second column shows the masses of the particles, where the most massive particle is the top quark with 173 GeV. There are particles with mass equal to zero, such as the gluons, photons and the neutrinos. \\
The third column corresponds to the charge of the particles. For the quarks, there are two different charges: $\frac{2}{3}$ for $u$,$c$ and $t$ and $-\frac{1}{3}$ for $d$,$s$ and $b$. For the leptons, the charge is negative, except for neutrinos, which charge is zero. 
The bosons have a integer charge; $\pm 1$ for $W$ and 0 for $Z$, $\gamma$, $H$ and $g$.	\\
The next column is the spin, where for fermions (quarks and leptons) is $\frac{1}{2}$ and for bosons is 1\footnote{In general, the fermions have non integer spin and the bosons have integer spin}. The Higgs boson is the only boson with spin 0.\\
The next one is the lifetime of the particle, where it shows how long the particles exist before decay. Stable lifetime means that the particles do not decay or can exist until they can interact with detectors.\\
The last column is the distance of particles moving with speeds near light speed. This column visualize the distance traveled in the predicted lifetime of the particle. And also give us a perspective of how hard is the detection of the particles due to lifetime.  \\

\begin{table}[!htbp]
\caption{Table of particles in SM, where it shows the masses, charges,spin, lifetime and distance traveled in a lifetime . Stable means no decay, or lifetime almost infinite\cite{pd}. Particles speeds at c=0.998}
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|c|c|c|c|c|p{2.5cm}|}
\hline 
Particle	& Mass (MeV/$c^2$) &Charge & spin &Lifetime (s) & Distance in lifetime (meters) \\ 
	\hline 
Up ($u$)	& 2.2 & $\frac{2}{3}$ & $\frac{1}{2}$ & stable & -\\ 
	\hline 
Charm ($c$)	& 1280 &$\frac{2}{3}$ & $\frac{1}{2}$ & $ 1.1 \times 10^{-12}$ & 5.21$\times 10^{-3}$ \\ 
	\hline 
Top	($t$)& 173100 & $\frac{2}{3}$ & $\frac{1}{2}$ & 5$\times$10$^{-25}$ &2.37$\times 10^{-15}$  \\ 
	\hline 
Down ($d$)	& 4.6 &$-\frac{1}{3}$ & $\frac{1}{2}$ & Stable & - \\ 
	\hline 
Strange ($s$)	& 96 &$-\frac{1}{3}$ & $\frac{1}{2}$ &$1.24 \times 10^{-8}$ & 58.7 \\ 
	\hline

Bottom ($b$)	& 4180 &$-\frac{1}{3}$ & $\frac{1}{2}$ &$1.3 \times 10^{-12}$  & 6.16 $\times 10^{-3}$\\ 
	\hline 
$W$ 	& 80379 &$\pm$1 & 1 & 3$\times$ 10$^{-25}$ & 1.42$\times 10^{-15}$\\ 
	\hline 
$Z$ & 91187.6 &0 & 1 & 3$\times$ 10$^{-25}$ &1.42$\times 10^{-15}$ \\ 
\hline
Photon ($\gamma$) & 0 &0 & 1&Stable & - \\ 
\hline
Gluon ($g$)	& 0 &0 & 1&Stable & - \\ 
	\hline 
Higgs ($H$)	& 125.18 &0 & 0& 1.56 $\times$ $10^{-22}$ & 7.39$\times 10^{-13}$ \\ 
	\hline 
Electron ($e$)& 0.511 & -1 &  $\frac{1}{2}$& Stable & - \\ 
	\hline 
Muon ($\mu$)	& 105.7 & -1 & $\frac{1}{2}$ & 2.2$\times$10$^{-6}$ & 10419.85 \\ 
	\hline 
Tau ($\tau$)	& 1776.86 &-1 & $\frac{1}{2}$ & 2.9$\times$10$^{-13}$ & 1.37$\times 10^{-3}$\\ 
	\hline 
$\nu_e$	$\nu_\mu$ $\nu_\tau$& 0 & 0 & $\frac{1}{2}$ & Stable & -\\
	\hline 
\end{tabular} 
\label{SM table}
\end{table}
\newpage

@@

The theory also explains Higgs bosons give the mass to the particles via Higgs Field and the mass of Higgs boson itself. The Higgs boson was discovered recently on year 2012 in the experiments ATLAS and CMS at CERN. The experiment was a proton proton collision with center of mass energy of 7 and 8 TeV,where investigators worked on the data and detecting various decay modes that generates final states such as WW, $\gamma \gamma$, ZZ, b$\bar{b}$ and $\tau \tau$ \cite{higgsd}. \\
The mass of this boson has been measured and its production and decay
rates are consistent with the SM prediction. The mass of the Higgs boson measured is 
approximately 125 MeV. The Higgs boson lifetime is 1.56 $\times$ $10^{-22}$. That is why the Higgs boson was very difficult to detect in the particle detectors. \\

@@

\section{Top quark}
The top quark is the most massive particle with a charge of 2/3 with spin of 1/2. First proposed by M. Kobayashi and T. Maskawa in 1973 to explain observed CP violations in Kaon decay\cite{griff}. On 1995, in the Tevatron at Fermilab in Illinois, the top quark was discovered via proton antiproton collisions with center of mass energy between 1.8 to 1.96 TeV in the CDF and D0 experiments.The top quark production process was via ggF, which generated a pair of $t\bar{t}$\cite{top}.
Since then,the properties of the top quark have been studied at the Tevatron Collider at Fermilab, and by the ATLAS and CMS experiments at the Large Hadron Collider (LHC) at CERN.\\
However, the top quark often receives special attention in new physics models because its mass requires near unity coupling to the Higgs boson\cite{top}.
As shown in table \ref{SM table}, the top quark lifetime is so small that they decay before they can hadronise, that is to say, the process to create hadrons from quarks and gluons.\\

@@

$W^i$ (i=1,2,3) are three SU(2) gauge bosons and B is a U(1) gauge boson.\footnote{$\epsilon^{ijk}$ represents the Levi-Civita tensor, where $\epsilon^{ijk}= \begin{dcases}
	+1 \quad \text{for even permutation} \\
	-1 \quad \text{for odd permutation}\\
	0 \quad \text{i=j, or j=k, or k=i } 
	\end{dcases}$.} In the end, those gauge bosons, after introduce the Higgs mechanism, become the $W^\pm$, $Z$ and $\gamma$ vector boson
\cite{ew1}\cite{ew2}. %Z bosons and photons will be formed by a combination of W and B fields.\\
\\

The second term refers to kinematic energy of fermions 
\begin{align}
\mathcal{L}_f=\sum_{m=1}^{3} \left(\bar{q}^m_{L}i \cancel{D}_L q^m_{L}+ 
\bar{l}^m_{L}i \cancel{D}_L l^m_{L} +\bar{u}_{R}^m i \cancel{D}_R u_{R}^m
+\bar{d}_{R}^m i \cancel{D}_R d_{R}^m+\bar{e}_{R}^m i \cancel{D}_R e_{R}^m \right)
\end{align}
where m is the family index, and L(R)
refer to the left (right) handed particles.\footnote{$\bar{\psi}$ denotes the adjoint spinor $\bar{\psi}=\psi^\dagger \gamma^0$.
$\cancel{D}=\gamma^\mu D_\mu$. $\gamma$ are the Dirac matrices}. 
The gauge covariant derivatives are given by
\begin{align}
{D}_L =\left(\partial_\mu+ig\sigma^iW^i_\mu +ig'Y B_\mu\right)\\
{D}_R =\left(\partial_\mu +ig' Y B_\mu\right)
\end{align}
where $\sigma^i$ are the Pauli matrices. L and R refers to left and right handed fields.
The left handed fields for the first family are defined by the following doublets
\begin{align}
  q_L=\left(\begin{array}{c}
 u_L \\
d_L
\end{array} \right) \qquad  l_L=\left(\begin{array}{c}
 \nu_{eL} \\
e^-_L
\end{array} \right)
\end{align}
and the right handed fields
\begin{align}
  u_R,d_R,e^-_R 
\end{align}

@@

In particle physics, cross section ($\sigma$) describes the likelihood of two particles interacting under certain conditions.
 Cross sections are expressed in barns, where 1 barn=$10^{-34}$ cm$^{2}$.
\\
The cross section is important in the evaluation of events for specific processes. In order to get the number of events for a specific process, the reaction rate $N$ is determined by the total cross section $\sigma$ and the luminosity L.\\
%L is called luminosity, that is the number of particle interactions can be produced in a detector per second. 
The unit of measurement of luminosity is cm$^{-2}$s$^{-1}$. Therefore, The reaction rate is

@@

In the table \ref{crt} shows the Higgs production cross section for pp collision and the number of events produced using a integrated luminosity of $35.9 fb^{-1}$, that is the luminosity measured for the Run 2 from LHC on 2016\cite{pd}.

@@

Here we can see ggF process have the biggest cross section of all Higgs mechanism and generates the greatest number of events possibles. For the processes tH and $t\bar{t}$H, they have the smallest cross section, so the probability for the processes is low and generates a small number of events. 
\pagebreak
\footnotetext{Taken from \url{ https://twiki.cern.ch/twiki/bin/view/LHCPhysics/LHCHXSWGCrossSectionsFigures}}
\section{Higgs decay rates}
In particle physics,two importatns properties are the lifetime and the decay rates of a particle. The lifetime is related to the total decay rate $\Gamma$, that is the probability per unit of time a particle decay, 
\begin{align}
  dN=-\Gamma N dt
\end{align}
with N the original number of particles before the decay. From here is evident that number of particles left in a decay is

@@

The number of particles decreases over time exponentially and this can be measured. With this, it is possible determinate the mean lifetime\cite{griff}
\begin{align}
\tau=\frac{1}{\Gamma}
\end{align}
In case of the decay rate to a specific process, it is necessary get the branching fraction of the process. 
The branching ratio for a decay process is the ratio of the number of particles which decay via a specific decay mode with respect to the total number of particles which decay via all decay modes.
\begin{align}
BR_i =\frac{\Gamma_i}{\sum_{i}\Gamma_i}
\end{align}
Where $\Gamma=\sum_i\Gamma_i$ is the total decay width (sum of all partial widths) of the particle and is related to lifetime of the particle: $\Gamma=1/\tau$.
Since the dimension of $\Gamma$ is the inverse of time, in our system of natural units, it is measured in inverse seconds,it has the same dimension as mass (or energy)\footnote{Cleaves H.J. (2011) Branching Ratio. In: Gargaud M. et al. (eds) Encyclopedia of Astrobiology. Springer, Berlin, Heidelberg}. The lifetime of the Higgs boson is predicted to be 1.56 $\times$ $10^{-22}$ seconds and corresponds to $\Gamma_H$ is about 4 MeV, but has not yet been measured due to the detector resolution\cite{cms-manual}.
\\
\begin{table}[!htbp] 
\caption[SM Higgs boson branching ratios for $M_H$ =125 GeV]{SM Higgs boson branching ratios for $M_H$ =125 GeV \protect \cite{pd}}
\centering
\begin{tabular}{|c|c|}
\hline
Higgs decay & Branching ratio (BR)\\
\hline
$H \rightarrow$ b$\bar{b}$ &$58.4\%$ \\
\hline
 $H \rightarrow$ $W^+W^-$ &$21.4\%$ \\
\hline
$H \rightarrow$ $\tau^+ \tau^-$ & $6.27\%$\\
\hline
$H \rightarrow$ ZZ &$2.62\%$\\
\hline
$H \rightarrow$ $\gamma\gamma$ &$0.227\%$\\
\hline
$H \rightarrow$ Z$\gamma$ &$0.153\%$\\
\hline
$H \rightarrow$ $\mu^+\mu^-$ &$0.0218\%$\\
\hline
\end{tabular}
\label{higgs1}
\end{table}
The decay $H \rightarrow$ $b\bar{b}$ have a branching ratio a bit more than 50$\%$, therefore is the most likely decay Higgs boson can decay. Due to this, $H\rightarrow b\bar{b}$ is a very studied process and most likely to detect a Higgs boson in the data after reconstruction. The decay $H \rightarrow$ $\mu^+\mu^-$ has a very low branching ratio. This decay is very rare but not impossible to detect in the future. In this work the majority of the Higgs decays detected are from $WW$, $ZZ$, $\tau \tau$\\

>>>>

In the recent years, there have been many studies in the Higgs production processes at CERN in pp collisions, such as gluon fusion (ggF), VBF, WH, ZH and $t\bar{t}$H and people have obtained consistent results in the search of Higgs in those processes. 
In this work, we will analyze the production of Higgs boson in association with a
single top quark (tH) in proton-proton collisions with the CMS experiment of the LHC. 
\\

The motivation of the work is this mechanism of production of the Higgs boson has not been observed before by any experiment. The exploration of Higgs production on the tH channel is subject relatively new. %The SM is used to study the tH channel and compare it with other Higgs production mechanisms results.
Many channels are yet out of
reach experimentally. At the same time, many theoretical
proposals are yet to be put on test. It is possible that studies on the Higgs boson give evidence of small deviations from the SM and test new physics that are beyond SM such as String Theory and Super-symmetry. \\
%********************************** % Third Section *************************************

@@

 Quarks combine to form composite particles called hadrons. Quarks cannot be found in isolation, they always exist in form of hadrons that can be mesons or baryons. Mesons, are bosons made up of a quark-antiquark pair, such as pion ($\pi$) and kaon ($K$). Mesons were theorized by Yukawa in 1937 and discovered in 1947 by Powell\cite{griff}.
Baryons are fermions made of 3 quarks. Examples of baryons are protons and neutrons. Both particles can interact via strong force. Since quarks and gluons only exist in bound, it is impossible to observe isolated quarks except the top quark, which decays before it has time to form a
bound state \cite{gross}.
\\

@@

Quarks have characteristics such as electric charge, mass, color charge, and spin. The quarks up, charm and top have in common a charge of $\frac{2}{3}$, while the down, strange and bottom quarks have charge of $-\frac{1}{3}$. Among them, the top quark is the most massive of all, while the up and down are the least massive. 
Some of the particles in the SM are not stable, so they decays and form particles with lower masses. Other particles do not decay or their lifetimes are very long. The final states can be represented with a Feynman diagram according to the SM vertices\cite{mark}.

@@

The last group of the family are the bosons, which have a spin 1 and 0. 
There are 5 bosons: $W^{\pm}$, $Z$ with neutral charge are responsible for weak force; photon, a massless particle, is responsible for electromagnetic force; gluon which also massless, is related to strong force and Higgs boson, recently discovered, is in charge of give the property of mass to particles via Higgs mechanism. Bosons let other particles to interact, even themselves, which is shown in the figure \ref{sm}. 
The $W$,$Z$ and Higgs bosons, can interact with the most of the particles, while the gluon can only interact with quarks and itself.
Those bosons act as mediators in decays and creation of complex
particles, as mesons and baryons.\\

@@

There are experiments that detect particles around the world, such as the large hadron collider (LHC) in Europe, the Fermilab in United States of America, KEK in Japan, etc. More information about the LHC will be described in the next chapters.
\\

The Table \ref{SM table} contains the characteristics of each particle in the SM. The second column shows the masses of the particles, where the most massive particle is the top quark with 173 GeV. There are particles with mass equal to zero, such as the gluons, photons and the neutrinos. 
The third column corresponds to the charge of the particles. For the quarks, there are two different charges: $\frac{2}{3}$ for $u$,$c$ and $t$ and $-\frac{1}{3}$ for $d$,$s$ and $b$. For the leptons, the charge is negative, except for neutrinos, which charge is zero. 
The bosons have a integer charge; $\pm 1$ for $W$ and 0 for $Z$, $\gamma$, $H$ and $g$.	
The next column is the spin, where for fermions (quarks and leptons) is $\frac{1}{2}$ and for bosons is 1\footnote{In general, the fermions have non integer spin and the bosons have integer spin}. The Higgs boson is the only boson with spin 0.
The next one is the lifetime of the particle, where it shows how long the particles exist before decay. Stable lifetime means that the particles do not decay or can exist until they can interact with detectors.
The last column is the distance of particles moving with speeds near light speed. This column visualize the distance traveled in the predicted lifetime of the particle. And also give us a perspective of how hard is the detection of the particles due to lifetime.  \\

\begin{table}[!htbp]
\caption[Table of particles in SM]{SM particle properties: mass, charges, spin, lifetime and distance traveled in a lifetime (assuming speed of 0.998*c). Stable means no decay, or lifetime almost infinite\cite{pd}.}
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|c|c|c|c|c|p{2.5cm}|}
\hline 
Particle	& Mass (MeV/$c^2$) &Charge & spin &Lifetime (s) & Distance (meters) \\ 
	\hline 
Up ($u$)	& 2.2 & $\frac{2}{3}$ & $\frac{1}{2}$ & stable & -\\ 
	\hline 
Charm ($c$)	& 1280 &$\frac{2}{3}$ & $\frac{1}{2}$ & $ 1.1 \times 10^{-12}$ & 5.21$\times 10^{-3}$ \\ 
	\hline 
Top	($t$)& 173100 & $\frac{2}{3}$ & $\frac{1}{2}$ & 5$\times$10$^{-25}$ &2.37$\times 10^{-15}$  \\ 
	\hline 
Down ($d$)	& 4.6 &$-\frac{1}{3}$ & $\frac{1}{2}$ & Stable & - \\ 
	\hline 
Strange ($s$)	& 96 &$-\frac{1}{3}$ & $\frac{1}{2}$ &$1.24 \times 10^{-8}$ & 58.7 \\ 
	\hline

Bottom ($b$)	& 4180 &$-\frac{1}{3}$ & $\frac{1}{2}$ &$1.3 \times 10^{-12}$  & 6.16 $\times 10^{-3}$\\ 
	\hline 
$W$ 	& 80379 &$\pm$1 & 1 & 3$\times$ 10$^{-25}$ & 1.42$\times 10^{-15}$\\ 
	\hline 
$Z$ & 91187.6 &0 & 1 & 3$\times$ 10$^{-25}$ &1.42$\times 10^{-15}$ \\ 
\hline
Photon ($\gamma$) & 0 &0 & 1&Stable & - \\ 
\hline
Gluon ($g$)	& 0 &0 & 1&Stable & - \\ 
	\hline 
Higgs ($H$)	& 125.18 &0 & 0& 1.56 $\times$ $10^{-22}$ & 7.39$\times 10^{-13}$ \\ 
	\hline 
Electron ($e$)& 0.511 & -1 &  $\frac{1}{2}$& Stable & - \\ 
	\hline 
Muon ($\mu$)	& 105.7 & -1 & $\frac{1}{2}$ & 2.2$\times$10$^{-6}$ & 10419.85 \\ 
	\hline 
Tau ($\tau$)	& 1776.86 &-1 & $\frac{1}{2}$ & 2.9$\times$10$^{-13}$ & 1.37$\times 10^{-3}$\\ 
	\hline 
$\nu_e$, $\nu_\mu$, $\nu_\tau$& 0 & 0 & $\frac{1}{2}$ & Stable & -\\
	\hline 
\end{tabular} 
\label{SM table}
\end{table}
\newpage

@@

The theory also explains Higgs bosons give the mass to the particles via Higgs Field and the mass of Higgs boson itself. The Higgs boson was discovered recently on year 2012 in the experiments ATLAS and CMS at CERN. The experiment was a proton proton collision with center of mass energy of 7 and 8 TeV, where investigators worked on the data and detecting various decay modes that generates final states such as WW, $\gamma \gamma$, ZZ, b$\bar{b}$ and $\tau \tau$ \cite{higgsd}. \\
The mass of this boson has been measured and its production and decay
rates are consistent with the SM prediction. The mass of the Higgs boson measured is 
approximately 125 MeV. The Higgs boson lifetime is 1.56 $\times$ $10^{-22}$. That is why the Higgs boson was very difficult to detect in the particle detectors. \\

@@

\section{Top quark}
The top quark is the most massive particle with a charge of 2/3 with spin of 1/2. First proposed by M. Kobayashi and T. Maskawa in 1973 to explain observed CP violations in Kaon decay\cite{griff}. On 1995, in the Tevatron at Fermilab in Illinois, the top quark was discovered via proton antiproton collisions with center of mass energy between 1.8 to 1.96 TeV in the CDF and D0 experiments.The top quark production process was via ggF, which generated a pair of $t\bar{t}$\cite{top}.
Since then, the properties of the top quark have been studied at the Tevatron Collider at Fermilab, and by the ATLAS and CMS experiments at the Large Hadron Collider (LHC) at CERN.
However, the top quark often receives special attention in new physics models because its mass requires near unity coupling to the Higgs boson\cite{top}.
As shown in Table \ref{SM table}, the top quark lifetime is so small that they decay before they can hadronise, that is to say, the process to create hadrons from quarks and gluons.\\

@@

$W^i$ (i=1,2,3) are three SU(2) gauge bosons and B is a U(1) gauge boson.\footnote{$\epsilon^{ijk}$ represents the Levi-Civita tensor, where $\epsilon^{ijk}= \begin{dcases}
	+1 \quad \text{for even permutation} \\
	-1 \quad \text{for odd permutation}\\
	0 \quad \text{i=j, or j=k, or k=i } 
	\end{dcases}$.} In the end, those gauge bosons, after introduce the Higgs mechanism, become the $W^\pm$, $Z$ and $\gamma$ vector boson
\cite{ew1,ew2}. %Z bosons and photons will be formed by a combination of W and B fields.\\
\\

The second term refers to kinematic energy of fermions 
\begin{align}
\mathcal{L}_f=\sum_{m=1}^{3} \left(\bar{q}^m_{L}i \cancel{D}_L q^m_{L}+ 
\bar{l}^m_{L}i \cancel{D}_L l^m_{L} +\bar{u}_{R}^m i \cancel{D}_R u_{R}^m
+\bar{d}_{R}^m i \cancel{D}_R d_{R}^m+\bar{e}_{R}^m i \cancel{D}_R e_{R}^m \right)
\end{align}
where m is the family index, and L(R)
refer to the left (right) handed particles.\footnote{$\bar{\psi}$ denotes the adjoint spinor $\bar{\psi}=\psi^\dagger \gamma^0$.
$\cancel{D}=\gamma^\mu D_\mu$. $\gamma$ are the Dirac matrices}. 
The gauge covariant derivatives are given by
\begin{align}
{D}_L =\left(\partial_\mu+ig\sigma^iW^i_\mu +ig'Y B_\mu\right)\\
{D}_R =\left(\partial_\mu +ig' Y B_\mu\right)
\end{align}
where $\sigma^i$ are the Pauli matrices. L and R refers to left and right handed fields.
The left handed fields for the first family are defined by the following doublets
\begin{align}
  q_L=\left(\begin{array}{c}
 u_L \\
d_L
\end{array} \right) \qquad  l_L=\left(\begin{array}{c}
 \nu_{eL} \\
e^-_L
\end{array} \right)
\end{align}
and the right handed fields
\begin{align}
  u_R, d_R, e^-_R 
\end{align}

@@

In particle physics, cross section ($\sigma$) describes the likelihood of two particles interacting under certain conditions.
 Cross sections are expressed in barns, where 1 barn=$10^{-34}$ cm$^{2}$.
\\
The cross section is important in the evaluation of events for specific processes. In order to get the number of events for a specific process, the reaction rate $N$ is determined by the total cross section $\sigma$ and the luminosity L.
%L is called luminosity, that is the number of particle interactions can be produced in a detector per second. 
The unit of measurement of luminosity is cm$^{-2}$s$^{-1}$. Therefore, The reaction rate is

@@

In the Table \ref{crt} shows the Higgs production cross section for pp collision and the number of events produced using a integrated luminosity of $35.9 fb^{-1}$, that is the luminosity measured for the Run 2 from LHC on 2016\cite{pd}.

@@

Here we can see ggF process have the biggest cross section of all Higgs mechanism and generates the greatest number of events possibles. For the processes tH and $t\bar{t}$H, they have the smallest cross section, so the probability for the processes is low and generates a small number of events. 
\pagebreak
\footnotetext{Taken from \url{ https://twiki.cern.ch/twiki/bin/view/LHCPhysics/LHCHXSWGCrossSectionsFigures}}
\section{Higgs decay rates}
In particle physics, two important properties are the lifetime and the decay rates of a particle. The lifetime is related to the total decay rate $\Gamma$, that is the probability per unit of time a particle decay, 
\begin{align}
  dN=-\Gamma N dt
\end{align}
with N the original number of particles before the decay. From here is evident that number of particles left in a decay is

@@

The number of particles decreases over time exponentially and this can be measured. With this, it is possible determinate the mean lifetime\cite{griff}
\begin{align}
\tau=\frac{1}{\Gamma}
\end{align}
In case of the decay rate to a specific process, it is necessary get the branching fraction of the process. 
The branching ratio for a decay process is the ratio of the number of particles which decay via a specific decay mode with respect to the total number of particles which decay via all decay modes.
\begin{align}
BR_i =\frac{\Gamma_i}{\sum_{i}\Gamma_i}
\end{align}
Where $\Gamma=\sum_i\Gamma_i$ is the total decay width (sum of all partial widths) of the particle and is related to lifetime of the particle: $\Gamma=1/\tau$.
Since the dimension of $\Gamma$ is the inverse of time, in our system of natural units, it is measured in inverse seconds, it has the same dimension as mass (or energy)\footnote{Cleaves H.J. (2011) Branching Ratio. In: Gargaud M. et al. (eds) Encyclopedia of Astrobiology. Springer, Berlin, Heidelberg}. The lifetime of the Higgs boson is predicted to be 1.56 $\times$ $10^{-22}$ seconds and corresponds to $\Gamma_H$ is about 4 MeV, but has not yet been measured due to the detector resolution\cite{cms-manual}.
\\
\begin{table}[!htbp] 
\caption[SM Higgs boson branching ratios for $M_H$ =125 GeV]{SM Higgs boson branching ratios for $M_H$ =125 GeV \protect \cite{pd}}
\centering
\begin{tabular}{|c|c|}
\hline
Higgs decay & Branching ratio (BR)\\
\hline
$H \rightarrow$ b$\bar{b}$ &$58.4\%$ \\
\hline
 $H \rightarrow$ $W^+W^-$ &$21.4\%$ \\
\hline
$H \rightarrow$ $\tau^+ \tau^-$ & $6.27\%$\\
\hline
$H \rightarrow$ ZZ &$2.62\%$\\
\hline
$H \rightarrow$ $\gamma\gamma$ &$0.227\%$\\
\hline
$H \rightarrow$ Z$\gamma$ &$0.153\%$\\
\hline
$H \rightarrow$ $\mu^+\mu^-$ &$0.0218\%$\\
\hline
\end{tabular}
\label{higgs1}
\end{table}
The decay $H \rightarrow$ $b\bar{b}$ have a branching ratio a bit more than 50$\%$, therefore is the most likely decay Higgs boson can decay. Due to this, $H\rightarrow b\bar{b}$ is a very studied process and most likely to detect a Higgs boson in the data after reconstruction. The decay $H \rightarrow$ $\mu^+\mu^-$ has a very low branching ratio. This decay is very rare but not impossible to detect in the future. In this work the majority of the Higgs decays detected are from $WW$, $ZZ$, $\tau \tau$\\
